---
title: "Customer Segmentation App"
author: "ZackM"
date: "6/19/2017"
output: slidy_presentation
runtime: shiny

<P style="page-break-before: always">
---


## <span style="color:navy">About The App</span><br /> 
<br />

###<span style="color:gray">__Visualization Objective:__</span>
###Provide business intelligence on customer segments to sales & marketing dept.<br />
<br />  
    

###<span style="color:gray">__Segments customers by:__</span> 

* ###*State*   
* ###*Age group*  
* ###*Average Revenue Per User* (ARPU)  
* ###*Recency*: Time in days between current date and date of most recent purchase--
* ###*Frequency*:  Number of purchases over the quarter--
* ###*Monetary Value*:  Total money spent by a customer over the quarter
<br />

---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
mydat2 <- read.csv("mydat2.csv", header = T, stringsAsFactors = F)
mydata <- read.csv("rfm_Final.csv", header = T, stringsAsFactors = F)
```

##<span style="color:navy">About The Data Set: Emerson Organics Sales, Q1 2017</span>

<h3><b><span style="color:gray">Raw Data:</span></b></h3><h4>&nbsp;&gt;&nbsp;25k+ transactions by ~14k customers across 18 states</h4>
```{r Table1,echo=FALSE, message=FALSE, warning=FALSE}

mydata <- read.csv("rfm_Final.csv", header = T, stringsAsFactors = F)
library(data.table)
library(tables)
library(dplyr)
library(knitr)
library(kableExtra)
library(xtable)
colnames(mydata)[6] <- "Name"
mydatsample <- head(mydata,2)

mydatsam1 <- mydatsample[,2:8]

#dtsam1 <- format(mydatsample1, justification = "c")
kable(mydatsam1, "html") %>%
  kable_styling(bootstrap_options = "striped", font_size =15, full_width = F, position = "left")
#print(mydatsam1, size="\\fontsize{2pt}{9pt}\\selectfont")

```
<br>

<h3><b><span style="color:gray">Transformed Data:</span></b></h3>
####> To calculate Recency, Freq., and MonetaryVal.
```{r Table2,echo=FALSE, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(sqldf)
library(data.table)
library(webshot)
library(shinyBS)
library(shinythemes)
library(shinydashboard)


mydata <- read.csv("rfm_Final.csv", header = T, stringsAsFactors = F)
sdata <- head(mydata[23:26,2:8])
rfm_dat <- mydata[,c(1,2,3,4,5,6,11,12,13)]
rfm_dat_unique <- rfm_dat[!duplicated(rfm_dat$cust_id),]
colnames(rfm_dat_unique)[6] <- "Name"
colnames(rfm_dat_unique)[7] <- "Recency"
colnames(rfm_dat_unique)[8] <- "Freq."
colnames(rfm_dat_unique)[9] <- "MonetaryVal($)"
df_brush <- rfm_dat_unique[,c(6,5,7,8,9)]
df <- rfm_dat_unique[,c(7,8,9)]
mydata <- data.table(mydata)
mydata <- mydata[, AgeBand:=(ifelse(mydata$age < 30, 1, ifelse(mydata$age < 40, 2, ifelse(mydata$age < 50, 3, ifelse(mydata$age < 60, 4,5)))))]
mydata$AgeGroup <- ifelse(mydata$AgeBand==1, "20s", ifelse(mydata$AgeBand==2, "30s", ifelse(mydata$AgeBand==3, "40s", ifelse(mydata$AgeBand==4, "50s", "60 & Above"))))
mydata$grand_total <- as.numeric(mydata$grand_total)
mydata2 <- sqldf("SELECT COUNT(DISTINCT(cust_id)) Cust_Count, 
                 SUM(grand_total) Revenue, 
                 ROUND(SUM(grand_total)/COUNT(DISTINCT cust_id), 2) ARPU,  
                 ROUND(AVG(Freq),2) AvgFreq, AgeGroup, state 
                 FROM mydata
                 GROUP BY state, AgeGroup")

mydatsam2 <- rfm_dat_unique[1:2, c(2,4,5,6,7,8,9)]
kable(mydatsam2, "html") %>%
  kable_styling(bootstrap_options = "striped", font_size =15, full_width = F, position = "left")
```

####> To calculate Avg.Freq. and ARPU by Age group:
```{r Table3,echo=FALSE, message=FALSE, warning=FALSE}
mydata2 <- data.frame(mydata2)
mydatsam3 <- head(mydata2,2)
kable(mydatsam3, "html") %>%
  kable_styling(bootstrap_options = "striped", font_size =15, full_width = F, position = "left")
```





## <span style="color:navy">App Dashboard: Plotting Customer Segments</span>


```{r, echo=FALSE,message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(sqldf)
library(data.table)
library(webshot)
library(shinyBS)
library(shinythemes)
library(shinydashboard)

mydata <- read.csv("rfm_Final.csv", header = T, stringsAsFactors = F)
sdata <- head(mydata[23:26,2:8])
rfm_dat <- mydata[,c(1,2,3,4,5,6,11,12,13)]
rfm_dat_unique <- rfm_dat[!duplicated(rfm_dat$cust_id),]
colnames(rfm_dat_unique)[6] <- "Name"
colnames(rfm_dat_unique)[7] <- "Recency"
colnames(rfm_dat_unique)[8] <- "Freq."
colnames(rfm_dat_unique)[9] <- "MonetaryVal($)"
df_brush <- rfm_dat_unique[,c(6,5,7,8,9)]
df <- rfm_dat_unique[,c(7,8,9)]
mydata <- data.table(mydata)
mydata <- mydata[, AgeBand:=(ifelse(mydata$age < 30, 1, ifelse(mydata$age < 40, 2, ifelse(mydata$age < 50, 3, ifelse(mydata$age < 60, 4,5)))))]
mydata$AgeGroup <- ifelse(mydata$AgeBand==1, "20s", ifelse(mydata$AgeBand==2, "30s", ifelse(mydata$AgeBand==3, "40s", ifelse(mydata$AgeBand==4, "50s", "60 & Above"))))
mydata$grand_total <- as.numeric(mydata$grand_total)
mydata2 <- sqldf("SELECT COUNT(DISTINCT(cust_id)) Cust_Count, 
                 SUM(grand_total) Revenue, 
                 ROUND(SUM(grand_total)/COUNT(DISTINCT cust_id), 2) ARPU,  
                 ROUND(AVG(Freq),2) AvgFreq, AgeGroup, state 
                 FROM mydata
                 GROUP BY state, AgeGroup")
mydata2 <- data.frame(mydata2)
library(shiny)


shinyApp(
  ui <- fluidPage(theme=shinytheme("yeti"),
    tags$head(tags$style(HTML("
    @import url('http://fonts.googleapis.com/css?family=Trajan:400,500,700');
    @import url('http://fonts.googleapis.com/css?family=Helvetica:400,500,700');
                                
                                h1{
                                font-family: 'Trajan';
                                font-style: normal;
                                font-weight: 500;
                                line-height: 2.1;
                                color: #4cc552;
                                }

                                h2{
                                font-family: 'Trajan';
                                font-style: normal;
                                font-weight: 300;
                                line-height: 0.5;
                                color: #0059b3;
                                }

                              h3{
                              font-family: 'Helvetica';
                              font-style: normal;
                              font-size: 7px;
                              line-height: 0.5;
                              color: #00000;
                              }

                        
                                #body{
                                #background-color:#e6e6fa
                                #}
                            
                                "))
              
              
                ),
  
  headerPanel(
      fluidRow(
        column(12, img(height=59, width =59, src =('http://www.goo.gl/iaZ4Wo')), "Emerson Organics"),
        column(6, offset = 3 , h2("U.S. Market Customer Segmentation"))
                  )),
    pageWithSidebar(
    headerPanel(' '),
    #  mainPanel(
    sidebarPanel('', width=2),
    mainPanel(
        tabsetPanel( 
  
          tabPanel("Introduction",
                   span(htmlOutput("Instruction", inline = FALSE), style='color: dark gray; font-size:17px'), htmlOutput("Instruction2"),tableOutput("sdata")),
          tabPanel("K-means Clusters",
            pageWithSidebar(
              headerPanel(''),
              sidebarPanel(
                selectInput('xcol', 'X Variable', names(df),
                            selected = names(df)[[3]], width = "100%"),
                selectInput('ycol', 'Y Variable', names(df),
                            selected=names(df)[[2]], width = "100%"),
                numericInput('clusters', 'Cluster count', 4,
                             min = 1, max = 5, width = "100%"), width = "3"
              ),
              mainPanel(
                textOutput(h2('RFM K-means Clusters')),
                plotOutput('plot1', width = "90%", height = "320px", brush = 'plot_brush',hover='plot_hover'), 
                span(htmlOutput('Instruction4', inline = 'false'), style='color: dark gray; font-size:10px'),
                dataTableOutput('info')
              )
            )),
          
          tabPanel("Count by Age Group",
                   pageWithSidebar(
                     headerPanel(''),
                     sidebarPanel(
                          selectInput("State1", "Select State", 
                                      choices = c("CA", 
                                                  "CT",
                                                  "FL",
                                                  "IL",
                                                  "LA",
                                                  "MA",
                                                  "MD",
                                                  "MI",
                                                  "MN",
                                                  "MO",
                                                  "NJ",
                                                  "NY",
                                                  "OH",
                                                  "PA",
                                                  "TX",
                                                  "VA",
                                                  "WA",
                                                  "WI"),
                                      selected = "CA", width = "100%"), width = 2),
                          
                   mainPanel(
                     plotOutput("BarChart1", width = "90%", height = "300px"),
                     dataTableOutput('results1')
                     )
                   )),
          tabPanel("ARPU by Age Group",
                   pageWithSidebar(
                     headerPanel(''),
                     sidebarPanel(
                       selectInput("State2", "Select State", 
                                   choices = c("CA", 
                                               "CT",
                                               "FL",
                                               "IL",
                                               "LA",
                                               "MA",
                                               "MD",
                                               "MI",
                                               "MN",
                                               "MO",
                                               "NJ",
                                               "NY",
                                               "OH",
                                               "PA",
                                               "TX",
                                               "VA",
                                               "WA",
                                               "WI"),
                                   selected = "CA", width = "100%"), width = 2),
                   
                   mainPanel(
                   plotOutput("BarChart2", width = "90%", height = "300px"),
                   dataTableOutput('results2')
                   )
                   ))
          #tabPanel("Data Table",
                   #dataTableOutput("results"))
       
        )
      )
    )
  ),
  
# SERVER FUNCTION  
  server <- function(input, output, session){
    selectedData <- reactive({
      rfm_dat_unique[, c(input$xcol, input$ycol)]
    })
    
    clusters <- reactive({
      kmeans(selectedData(), input$clusters)
    })
    
    output$plot1 <- renderPlot({
      palette(c("#E41A1C", "#377EB8", "#4DAF4A", "#984EA3",
                "#FF7F00", "#FFFF33", "#A65628", "#F781BF", "#999999"))
      
      par(mar = c(5, 4, 0, 1))
      plot(selectedData(),
           col = clusters()$cluster,
           pch = 20, cex.axis = 1, cex.lab = 1.25)
      points(clusters()$centers, pch = 4, cex = 2.5 , lwd = 2)
    })
# Brushed Points - K Means    
    output$info <- renderDataTable({
      brushedPoints(df_brush, input$plot_brush, xvar = input$xcol, yvar = input$ycol)}, 
      options = list(scrollY = 200, lengthMenu = c(5,10,20,30)))
    
    filtered1 <- reactive({ 
      mydata2%>%filter(mydata2$state==input$State1)
    })
    
    output$results1 <- renderDataTable({
      as.data.frame(filtered1())}, 
      options = list(scrollY = 200, lengthMenu = c(5,10,20,30)))
    
    
    filtered2 <- reactive({
      mydata2%>%filter(mydata2$state==input$State2)
    })
    
    output$results2 <- renderDataTable({
      as.data.frame(filtered2())},
      options = list(scrollY = 200, lengthMenu = c(5,10,20,30)))
    
    
    output$Instruction <- renderText(paste("<br><b> &emsp; &nbsp; &nbsp; &nbsp; &nbsp; &emsp; Welcome to the Emerson Organics' Segmentation App<sup>&copy</sup>  !</b>"))
                        
    output$Instruction2 <- renderText(paste("<b>","<br>","<br>",">>> Segmentation: By State, By Age, By Revenue:","</b>","<br>","Customers may be segmented based on Age Group or <i>Avg. Revenue Per User </i>(ARPU)","<br>","Select a state from the dropdown in the left column","<br>","Click on the relevant tabs to generate the plots or data table","<br>","<br>","<b>>>> About The Dataset:","</b>","<br>","Quarterly sales database records the following variables from each transaction:","<br>","Customer ID, Order Date, Order ID, Region, Name, Age, and Grand Total($)","<br>","A sample of the same data (Q1,2017) is printed below for your perusal:","<br>","<br"))
    
    output$Instruction4 <- renderText(paste('Select Points On The ScatterPlot For Table Below'))
    
    output$sdata <- renderTable(sdata)
    
    
    
    output$BarChart1 <- renderPlot({
      
      ggplot(filtered1(), aes(AgeGroup)) + geom_bar(aes(weight = Cust_Count), width = 0.4) + theme(text = element_text(size = 14), axis.text = element_text(size=12)) + labs(x = "Age Group", y = "Total Customers")
    })
    
    output$BarChart2 <- renderPlot({
      
      ggplot(filtered2(), aes(AgeGroup)) + geom_bar(aes(weight = ARPU), width = 0.4) + theme(text = element_text(size = 14), axis.text = element_text(size=12)) + labs(x="Age Group", y = "Avg. Revenue Per User ($)")
    })
    
  },
options = list(height = 500)
)

   
```



